---
title: "si618hw1_report_wanjun"
author: "Jun Wang"
date: "2016/10/28"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## SI 618 Homework 1
## Step 1: Load data

First the provided TSV data file is loaded into R using the **read.table()** function. Here are the first 15 rows of the data frame:

```{r echo=FALSE}
setwd("C:\\UM\\Courses\\SI\\SI618\\HW\\HW1")
countrydata_withregion<-read.table("countrydata_withregion.tsv",quote = "\"",head=TRUE,sep="\t")
head(countrydata_withregion,15)
```

## Step 2: Scatter plot of log transformed data

Natural logarithms of the area and the population of each country are computed and used to produce the following scatter plot using the **plot()** function.

```{r echo=FALSE}
plot(log(countrydata_withregion$area),log(countrydata_withregion$population))
```

## Step 3: Data aggregation by region
The areas and populations of all countries in a region are summed up using the **aggregate()** function, respectively. Then the following two pie charts are created using the **pie()** function.

```{r echo=FALSE}
area_region<-aggregate(countrydata_withregion$area,by=list(countrydata_withregion$region),FUN=sum)
countrydata_withregion$population<-as.numeric(countrydata_withregion$population)
popu_region<-aggregate(countrydata_withregion$population,by=list(countrydata_withregion$region),FUN=sum)
pie(area_region$x,labels=area_region$Group.1,main = "Area of Regions")
pie(popu_region$x,labels=area_region$Group.1,main = "Population of Regions")
```

## Step 4: Visualization of Population per sq km of Regions
A new data frame is created to contain the population per sq km of each region using the **data.frame()** function. The data frame is then sorted by population per sq km in decreasing order with the help of the **order()** function. Finally, the following bar plot is created using the **barplot()** function.

```{r echo=FALSE}
popu_area<-popu_region$x/area_region$x
df_popu_area<-data.frame(popu_region$Group.1,popu_area)
df_ordered<-df_popu_area[order(df_popu_area$popu_area,decreasing = TRUE),]
barplot(df_ordered$popu_area,names.arg = df_ordered$popu_region.Group.1,xlab = "Population per sq km of Regions",cex.names = 0.45,cex.axis = 0.53,cex.lab=0.53)
```