---
title: "si618hw2report_wanjun"
author: "JW"
date: "2016/11/07"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#SI 618 Homework 2
##Load and Cleanse Data (5 points)

First load the TSV data file created in part 1 into R using the **read.table()** function. Then apply listwise deletion to remove 
records with missing data and display the first 15 rows of the data frame.

```{r,echo=FALSE}
library(ggplot2)
library(plyr)
setwd("C:\\UM\\Courses\\SI\\SI618\\HW\\HW2")
businessdata<-read.table("businessdata.tsv",sep = '\t',head=TRUE,quote = "",comment.char = "")
businessdata<-na.omit(businessdata)
head(businessdata,15)
```

##Histograms of Star Ratings (10 points)

Using the **qplot()** function, plot histograms of the distribution of star ratings, faceted by state, using a bin width of 0.5.

```{r, echo=FALSE}
qplot(x=stars,facets = .~state,fill=state,main = "Histograms of Star Ratings",binwidth=0.5,data=businessdata)
```

##Histograms of Review Counts (10 points)
Using the **qplot()** function, plot the distribution of review counts as a histogram. First, plot a histogram that simply uses the raw values in review_count:

```{r,echo=FALSE}
qplot(x=review_count,main="Histograms of Review Counts",xlab="Review Counts",data = businessdata)
```

Then for comparison, plot a histogram that transforms the review counts using the log() function.

```{r,echo=FALSE}
qplot(x=log(review_count),main="Histograms of Review Counts",xlab="log(Review Counts)",data = businessdata)
```

We can see the distribution of review counts has a long tail. To zoom in on the part of the distribution with review counts between 0 and 100, 
use the **subset()** function to select just the data with review count <= 100. And then plot the (non-log-transformed) histogram again using a 
bin width of 2, and also include a overlay that plots a probability density function over the review counts. (Recall that you can produce 
an overlay by passing a list of geometry names to the geom= argument in qplot().)

```{r,echo=FALSE}
subset_bus = subset(businessdata,review_count<=100&review_count>=0,review_count)
qplot(x=subset_bus$review_count,main = "Histograms of Review Counts (Filtered)",xlab="Review Counts",geom="density")+
  geom_histogram(aes(y=(..count..)/(2*sum(..count..))),center=0.99,binwidth = 2,alpha=0.5)
```

##Boxplot of Star Ratings by States (5 points)
Here, map the value of the state variable to the "color" and "fill" aesthetics, and plot a boxplot of the star ratings by state.

```{r,echo=FALSE}
qplot(state, stars, data = businessdata, geom = "boxplot", colour = state, fill = state, alpha = 0.5, main = "Star Ratings by States")+scale_alpha(guide = 'none')
```

##Jittered Plot of Star Ratings by State (5 points)
Here, map the value of the state variable to the "color" and "fill" aesthetics, and plot a jittered point plot of the star ratings by state.

```{r,echo=FALSE}
qplot(state, stars, data = businessdata, geom = "jitter", colour = state, fill = state, alpha = 0.5, main = "Star Ratings By States") + scale_alpha(guide = 'none')
```

##Bar Chart of Number of Businesses by State (5 points)
Plot a bar chart showing the number of businesses in the dataset for each state.

```{r,echo=FALSE}
qplot(state, data = businessdata)
```

##Scatterplot of Stars and Review Counts (5 points)
Plot a scatterplot showing the star rating and review count for all businesses, mapping the state variable to the color aesthetic.

```{r,echo=FALSE}
qplot(x=stars,y=review_count,data = businessdata,fill=state,color=state)
```

##Slice and Dice Data with plyr
###Subsetting Data (15 points)
We first rank the business in each city for each main category and display the first 20 results of the resulting list.

```{r,echo=FALSE}
ddply_bus<-ddply(businessdata,c("city","main_category"),transform,rank=rank(-stars,ties.method = "first"))
head(ddply_bus,20)
```

Then the top 5 ranked businesses in each city for each main category are found. We're interested in the top 5 businesses that have a main
category of "Thai" in each city. The data should be ordered by city name, and then by rank. The result is listed below.

```{r,echo=FALSE}
thai_bus<-subset(businessdata,main_category=="Thai")
ddply_thai<-ddply(thai_bus,.(city),transform,rank=rank(-stars,ties.method = "first"))
thai_top5_all = subset(ddply_thai,rank<=5)
thai_top5 = subset(ddply_thai[,c(2,1,7)], rank <= 5)
thai_top5<-arrange(thai_top5,city,rank)
thai_top5
```

###Summarize Data (5 points)
Compute the mean star ratings grouped by state, for all the selected top 5 Thai restaurants selected in the previous section. Plot the mean star rating using a point chart, 
use the size of the point to indicate the mean review count for that state's restaurants, and sort the states by descending mean rating.

A very useful way to control the order in which the states are sorted in the plot is the **reorder()** function.

* The **reorder** function operates on a variable and takes three parameters:
    + The variable whose factors you want to reorder
    + A numeric vector of equal length (i.e. that has one value per possible value of the given variable. In this case, one value per possible state.)
    + A function to apply to the numeric vector (optional: the identity function if not specified)
  
Since R treats each unique state string like an unordered factor, you can use reorder() to impose an ordering on the states according to the numeric vector you provide.  
We'll cover factors more in an upcoming class.

```{r, echo=FALSE}
mean_stars = ddply(thai_top5_all,.(state),summarise,mean_star=mean(stars),mean_review=mean(review_count))
qplot(reorder(state,mean_star,function(x)-x),mean_star,data=mean_stars,geom = "point",ylab = "Mean Star Rating",xlab = "State",size=mean_review,main = "Mean Star Rating of Top Thai Restaurants, by State")
```
